Learning From Massive Noisy Labeled Data for Image Classification

### 2015-04-07 paper.tex
Modification by Yang Yi

One possible solution is to automatically collect large amount of annotations from the Internet web images~\cite{gong2014multiview} (i.e. extracting tags from the surrounding texts or keywords from search engines) and directly use them as ground truth to train deep models. Unfortunately, these labels are extremely unreliable due to the various types of noise (i.e. labeling mistakes from annotators or computing errors from extraction algorithms). Many works have shown that these noisy labels could adversely impact the classification accuracy of the induced classifiers~\cite{zhu2004class,nettleton2010study,pechenizkiy2006class}. Various noise-robust algorithms are developed but experiments show that performances of classifiers inferred by robust algorithms are still affected by label noise. Other noise cleaning algorithms are developed~\cite{barandela2000decontamination,brodley2011identifying}, but these approaches are difficult in distinguishing informative hard examples from harmful mislabeled ones.

@@

Although annotating all the noisy data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people first initialize the parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetune on the smaller dataset of specific task~\cite{oquab2013learning,azizpour2014generic,donahue2013decaf}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a Convolutional Neural Network (CNN) from scratch with limited clean labels and massive noisy labels is better than finetuning it only on the clean labels. Other approaches address the problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets. Therefore, it is inevitable to develop a better way of using huge amount of noisy labeled data.

Our goal is to build an end-to-end deep learning system that is capable of training with both limited clean labels and massive noisy labels more effectively. Figure~\ref{fig:overview} shows the framework of our approach. We collect a million clothes images from E-commerce websites. Each image is automatically assigned a noisy label according to the keywords in its surrounding text. After that, we manually refine 72409 image labels, which constitute a clean sub-dataset. All the data are then used to train CNNs, while the major challenge is to identify and correct the wrong labels during the training process.

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the ground truths and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modifies a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However, they assume noisy labels are conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit the real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. It shows that images tend to be mislabeled may share similar visual patterns. This is because different sellers have their own bias on different categories, thus they may provide wrong keywords for many similar clothes. Based on these observations, we introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

The proposed probabilistic model will build the relations among the image, noisy label, ground truth label, and the noise type, where the latter two are treated as latent variables. We use the Expectation-Maximization (EM) algorithm to solve the problem and integrate it into the training process of CNNs. Experiments on the collected dataset show that our model can better detect and correct the wrong labels.

Our contribution comes from three aspects. First, we study the causes of noisy labels in real-world data and describe all the variables with a novel probabilistic model. Second, we integrate the model into existing deep learning framework to help detect and correct wrong labels. Different training strategies are investigated to make the CNNs learn from better supervisions. Finally, a large-scale clothing dataset is collected with both noisy and clean labels, which will be released for academic use.

% TODO: combine these two figures
\begin{figure}[t]
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/noise_level.pdf}
\end{center}
\caption{The probability of an image suffering from different noise types depend on the image itself. Although all the images belong to class ``T-shirt'', the top two images can be easily recognized, while the bottom two tend to be confused with class ``Chiffon''.}
\label{fig:compare_noise_level}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/noise_cluster.pdf}
\end{center}
\caption{Images tend to be mislabeled usually share similar visual patterns.}
\label{fig:noise_cluster}
\end{figure}

% section related_work (end)

\section{Related Work} % (fold)
\label{sec:related_work}
For most of the related works including effect of label noises, taxonomy of label noises, robust algorithms and noise cleaning algorithms for learning with noisy data, we refer to \cite{frenay2013classification} for a comprehensive review.

{\bf Direct learning with noisy labels:} Among other consequences, many works have shown that noise can adversely impact the classification performances of induced classifiers \cite{zhu2004class}. Experiments in the literature show that the performances of classifiers inferred by label noise-robust algorithms are still affected by label noise. These methods seem to be adequate only for simple cases of label noise that can be safely managed by overfitting avoidance \cite{frenay2013classification}.

{\bf Semi-supervised learning:} Apart from modeling label noise explicitly, some semi-supervised learning algorithms are developed to utilize weakly labeled or even unlabeled data. Label Propagation method \cite{zhu2002learning} explicitly uses ground truths of well-labeled data to classify unlabeled samples. However, it suffers from computing pairwise distance, which has quadratic complexity with the number of data samples and cannot be applied on large-scale datasets. Chen \etal~\cite{chen2013fast} tackled this problem in the scenario of image tagging by a co-regularized duo classifier that can efficiently use both the image and incomplete tags. Weston \etal~\cite{weston2012deep} proposed to embed a pairwise loss in the middle layer of a deep neural network, which benefits the learning of discriminative features. But they need extra information about whether a pair of unlabeled images belong to the same class, which cannot be obtained in our problem.

{\bf Transfer learning:} The success of Convolutional Neural Networks (CNNs) lies in their capability of learning rich and hierarchical image features. However, the model parameters cannot be properly learned when training data is not enough. Researchers proposed to conquer this problem by first initializing the CNN parameters with a model pretrained on a larger yet related dataset, and then finetuning it on the smaller dataset of specific task \cite{krizhevsky2012imagenet, oquab2013learning, azizpour2014generic,donahue2013decaf}. Nevertheless, this transfer learning scheme could be suboptimal when the two tasks are just loosely related. In our case of clothing classification, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it from an ImageNet pretrained model only on the clean labels.

{\bf Noise modeling with deep learning:} Various methods have been proposed to handle label noise in different problem settings, but there are very few works about deep learning from noisy labels \cite{mnih2012learning, larsen1998design, sukhbaatar2014learning}. \cite{mnih2012learning} builds a simple noise model for aerial images but only considers binary classification. \cite{larsen1998design} assumes label noise is independent from the true class label which is a simple and a specific case. \cite{sukhbaatar2014learning} generalizes from them by considering multi-class classification and modeling class dependent noise, but they assume the noise is conditionally independent with the image content, ignoring the hardness of labeling images of different confusing levels.
% section related_work (end)

@@

After that we manually refine the noisy labels of a small portion of all the images and split them into training ($\mathcal{D}_c$), validation and test sets. On the other hand, the remaining data construct the noisy labeled training dataset ($\mathcal{D}_n$). A crucial step here is to remove from $\mathcal{D}_c$ and $\mathcal{D}_n$ the images that are near duplicate with some image in the validation and test sets, which ensures the reliability of our test protocol. Finally, the size of training datasets are $|\mathcal{D}_c| = 47570$ and $|\mathcal{D}_n| = 10^6$, while validation and test set have 14313 and 10526 images, respectively.

>>>>

One possible solution is to automatically collect large amount of annotations from the Internet web images~\cite{gong2014multiview} (i.e. extracting tags from the surrounding texts or keywords from search engines) and directly use them as ground truth to train deep models. Unfortunately, these labels are extremely unreliable due to the various types of noise (i.e. labeling mistakes from annotators or computing errors from extraction algorithms). Many works have shown that these noisy labels could adversely impact the classification accuracy of the induced classifiers~\cite{zhu2004class,nettleton2010study,pechenizkiy2006class}. Various label noise-robust algorithms are developed but experiments show that performances of classifiers inferred by robust algorithms are still affected by label noise~\cite{bartlett2006convexity,teng2001comparison}. Other data cleansing algorithms are proposed~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches are difficult in distinguishing informative hard examples from harmful mislabeled ones.

@@

Although annotating all the data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people first initialize the parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetune on the smaller dataset of specific tasks~\cite{oquab2013learning,azizpour2014generic,donahue2013decaf}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a Convolutional Neural Network (CNN) from scratch with limited clean labels and massive noisy labels is better than finetuning it only on the clean labels. Other approaches address the problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets. Therefore, it is inevitable to develop a better way of using huge amount of noisy labeled data.

Our goal is to build an end-to-end deep learning system that is capable of training with both limited clean labels and massive noisy labels more effectively. Figure~\ref{fig:overview} shows the framework of our approach. We collect 1,000,000 clothing images from E-commerce websites. Each image is automatically assigned with a noisy label according to the keywords in its surrounding text. After that, we manually refine 72,409 image labels, which constitute a clean sub-dataset. All the data are then used to train CNNs, while the major challenge is to identify and correct the wrong labels during the training process.

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the ground truths and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modifies a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumes noisy labels are conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit the real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. It shows that images tend to be mislabeled may share similar visual patterns. This is because different sellers have their own bias on different categories, thus they may provide wrong keywords for many similar clothes. Based on these observations, we introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

The proposed probabilistic model build the relations among the image, noisy label, ground truth label, and the noise type, where the latter two are treated as latent variables. We use the Expectation-Maximization (EM) algorithm to solve the problem and integrate it into the training process of CNNs. Experiments on the collected dataset show that our model can better detect and correct the wrong labels.

Our contribution comes from three aspects. First, we study the cause of noisy labels in real-world data and describe it with a novel probabilistic model. Second, we integrate the model into existing deep learning framework to help detect and correct wrong labels. We study different training strategies to make the CNNs learn from better supervisions. Finally, we collect a large-scale clothing dataset with both noisy and clean labels, which will be released for academic use.

% TODO: combine these two figures
\begin{figure}[t]
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/noise_level.pdf}
\end{center}
\caption{The probability of an image suffering from different noise types depends on the image itself. Although all the images belong to class ``T-shirt'', the top two images can be easily recognized, while the bottom two tend to be confused with class ``Chiffon''.}
\label{fig:compare_noise_level}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/noise_cluster.pdf}
\end{center}
\caption{Images tend to be mislabeled usually share similar visual patterns.}
\label{fig:noise_cluster}
\end{figure}
% section introduction (end)

\section{Related Work} % (fold)
\label{sec:related_work}
For most of the related works including effect of label noises, taxonomy of label noises, robust algorithms and noise cleaning algorithms for learning with noisy data, we refer to~\cite{frenay2013classification} for a comprehensive review.

{\bf Direct learning with noisy labels:} Many works have shown that label noise can adversely impact the classification accuracy of induced classifiers~\cite{zhu2004class}. To better handle label noise, some approaches rely on training classifiers with label noise-robust algorithms~\cite{beigman2009learning,manwani2013noise}. However,~\cite{bartlett2006convexity} proves that most of the loss functions are not completely robust to label noise. Experiments in~\cite{teng2001comparison} show that the classifiers inferred by label noise-robust algorithms are still affected by label noise. These methods seem to be adequate only when label noise can be safely managed by overfitting avoidance~\cite{frenay2013classification}. On the other hand, some label noise cleansing methods are proposed to remove or correct mislabeled instances~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches are difficult in distinguishing informative hard examples from harmful mislabeled ones. Thus they may remove too many instances and the overcleansing can reduce the performances of classifiers~\cite{matic1992computer}.

{\bf Semi-supervised learning:} Apart from directly learning with label noise, some semi-supervised learning algorithms are developed to utilize weakly labeled or even unlabeled data. Label Propagation method~\cite{zhu2002learning} explicitly uses ground truths of well-labeled data to classify unlabeled samples. However, it suffers from computing pairwise distance, which has quadratic complexity with the number of samples and cannot be applied on large-scale datasets. Chen \etal~\cite{chen2013fast} tackled this problem in the scenario of image tagging by a co-regularized duo classifier that can efficiently use both the image and incomplete tags. Weston \etal~\cite{weston2012deep} proposed to embed a pairwise loss in the middle layer of a deep neural network, which benefits the learning of discriminative features. But they need extra information about whether a pair of unlabeled images belong to the same class, which cannot be obtained in our problem.

{\bf Transfer learning:} The success of Convolutional Neural Networks (CNNs) lies in their capability of learning rich and hierarchical image features. However, the model parameters cannot be properly learned when training data is not enough. Researchers proposed to conquer this problem by first initializing the CNN parameters with a model pretrained on a larger yet related dataset, and then finetuning it on the smaller dataset of specific task~\cite{krizhevsky2012imagenet, oquab2013learning, azizpour2014generic,donahue2013decaf}. Nevertheless, this transfer learning scheme could be suboptimal when the two tasks are just loosely related. In our case of clothing classification, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it from an ImageNet pretrained model only on the clean labels.

{\bf Noise modeling with deep learning:} Various methods have been proposed to handle label noise in different problem settings, but there are very few works about deep learning from noisy labels~\cite{mnih2012learning, larsen1998design, sukhbaatar2014learning}.~\cite{mnih2012learning} builds a simple noise model for aerial images but only considers binary classification.~\cite{larsen1998design} assumes label noise is independent from the true class label which is a simple and a specific case.~\cite{sukhbaatar2014learning} generalizes from them by considering multi-class classification and modeling class dependent noise, but they assume the noise is conditionally independent with the image content, ignoring the hardness of labeling images of different confusing levels. Our work can be viewed as a generalization of~\cite{sukhbaatar2014learning,natarajan2013learning} and our model is flexible enough to not only class dependent noise but also input dependent noise.
% section related_work (end)

@@

After that we manually refine the noisy labels of a small portion of all the images and split them into training ($\mathcal{D}_c$), validation and test sets. On the other hand, the remaining data construct the noisy labeled training dataset ($\mathcal{D}_n$). A crucial step here is to remove from $\mathcal{D}_c$ and $\mathcal{D}_n$ the images that are near duplicate with some image in the validation and test sets, which ensures the reliability of our test protocol. Finally, the size of training datasets are $|\mathcal{D}_c| = 47570$ and $|\mathcal{D}_n| = 10^6$, while validation and test set have 14,313 and 10,526 images, respectively.
